---
title: "Reproducible Research Course Assignment 1"
author: "PK"
date: "April 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}


#####Installing required libraries

library(knitr)  
opts_chunk$set(echo = TRUE)  
library(dplyr)  
library(ggplot2)
```
```{r}
#####To change the date into dateformat, installing lubridate package  
library(lubridate)  
```

###Reading activity data .csv file  
```{r}
activitydata<-read.csv("activity.csv", header = TRUE, sep=',', colClasses = c("numeric", "character", "integer"))  
```
####Changing the date format  
```{r}
activitydata$date<-ymd(activitydata$date) 
```
####Mean of Total Number of Steps taken per day
```{r}
steps <- activitydata %>%  
  filter(!is.na(steps)) %>%  
  group_by(date) %>%  
  summarize(steps = sum(steps)) %>%  
  print  
```

####Plotting the Histogram of All Steps Taken per Day Using 'ggplot2' Library and   printing the mean and median of the steps
```{r}
ggplot(steps, aes(x = steps)) +
  geom_histogram(fill = "red", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")  
meanofsteps <- mean(steps$steps, na.rm = TRUE)  
medianofsteps <- median(steps$steps, na.rm = TRUE)  
print(meanofsteps)  
print(medianofsteps) 
```

####Summary of Average Daily Pattern of Steps
```{r}
interval <- activitydata %>%  
  filter(!is.na(steps)) %>%  
  group_by(interval) %>%  
  summarize(steps = mean(steps))  
```
####Plotting  A Time Series Plot (i.e. type = "l") of the 5-minute Interval (x-axis) and the Average Number of Steps Taken, Averaged Across All days (y-axis)
```{r}
ggplot(interval, aes(x=interval, y=steps)) +
  geom_line(color = "blue") 
```

#### Printing Summary of 5-minute Interval, On Average Across All the Days in the Dataset, Contains the Maximum Number of Steps
```{r}
interval[which.max(interval$steps),]  
```

####Impiuting the Missing Data
######1. Total No. of Missing Values in the Dataset
```{r}
print(sum(is.na(activitydata$steps)))  
```

######2. Creating a New Dataset That is Equal to the Original Dataset but with the Missing Data Filled in Using tapply Function
```{r}
alldata <- activitydata  
nas <- is.na(alldata$steps)  
avgofinterval <- tapply(alldata$steps, alldata$interval, mean, na.rm=TRUE, simplify=TRUE) 
alldata$steps[nas] <- avgofinterval[as.character(alldata$interval[nas])]  
print(summary(is.na(alldata$steps)))  
```

######3. Calculaing and Plotting Histogram for The Number of Steps Taken in Each 5-minute Interval per Day Using dplyr and group by interval. 
```{r}
allsteps <- alldata %>%  
  filter(!is.na(steps)) %>%  
  group_by(date) %>%  
  summarize(steps = sum(steps)) %>%  
  print
ggplot(allsteps, aes(x = steps)) +
  geom_histogram(fill = "pink", binwidth = 1000) +
  labs(title = "Histogram of all Steps per day with missing values", x = "Steps per day", y = "Frequency")  
```

######4. Calculating Meand and Median 
```{r}
meanofallsteps <- mean(allsteps$steps, na.rm = TRUE)  
medianofallsteps <- median(allsteps$steps, na.rm = TRUE)  
print(meanofallsteps)  
print(medianofallsteps)  
```

####Checking the Differences in Activity Patterns between Weekdays and Weekends 
```{r}
alldata <- mutate(alldata, weektype = ifelse(weekdays(alldata$date) == "Saturday" | weekdays(alldata$date) == "Sunday", "weekend", "weekday"))  
alldata$weektype <- as.factor(alldata$weektype)  
head(alldata)  
```
####Plotting the above Result Using ggplot2 Package
```{r}
allinterval <- alldata %>%  
  group_by(interval, weektype) %>%  
  summarise(steps = mean(steps))  
 s<-ggplot(allinterval, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)  
  print(s)  
```